% Add GRAMMAR REPOSITORY 
% http://code.google.com/p/grammar-repository/
%\bibitem{perl6}
%{\sc Allison~Randal, Dan~Sugalski, L.~T.} 2004.
%\newblock {\em \em Perl 6 and Parrot Essentials}.
%\newblock O'Reilly Media.

\bibitem{Johnsonyacc}
{\sc Johnson, S.~C.} 1979.
\newblock Yacc: Yet another compiler compiler.
\newblock {\em AT\&T Bell Laboratories Technical Report July 31, 1978\/}~{\em
  2}, 353--387.

\bibitem{bison}
{\sc Donnelly, C.} {\sc and} {\sc Stallman, R.~M.} 1995.
\newblock { Bison: the yacc-com\-pat\-i\-ble parser generator}.
\newblock Technical report, {F}ree {S}oftware {F}oundation, 675 Mass Ave,
  Cambridge, MA 02139, Tel: (617) 876-3296.

\bibitem{elkhound}
{\sc Mcpeak, S.} 2004.
\newblock Elkhound: A fast, practical GLR parser generator.
\newblock Available at: {\tt http://scottmcpeak.com/elkhound/}.

\bibitem{tomita2}
{\sc Tomita, M.} 1990.
\newblock The generalized LR parser/compiler 
\newblock {m Proceedings of International Conference on Computational
  Linguistics (COLING'90)}. Helsinki, Finland, 59--63.

\bibitem{kelbt}
{\sc Thurston, A.~D.} {\sc and} {\sc Cordy, J.~R.} 2006.
\newblock A backtracking LR algorithm for parsing ambiguous context-dependent
  languages.
\newblock In { 2006 Conference of the Centre for Advanced Studies on
  Collaborative Research (CASCON 2006)}. Toronto, 39--53.

\bibitem{passos}
{\sc Teixeira Passos, L. Bigonha, M. A.S.,  Bigonha, R.S.},
\newblock A Methodology for Removing LALR(k) Conflicts
\newblock { Journal of Universal Computer Science}~{13} 737--752 2007.

\bibitem{Rodriguez:Leon}
{\sc Rodr\'iguez-Le\'on C.},
\newblock {\tt Parse::Eyapp} Manuals. 2011.
\newblock \\CPAN: {\tt http://search.cpan.org/dist/Parse-Eyapp/ } 
\newblock \\google-code: {\tt http://code.google.com/p/parse-eyapp/ } 

\bibitem{Wall:Perl}
{\sc Wall~L., Christiansen~T., S.~R.} 1996.
\newblock {\em Programming Perl}.
\newblock O'Reilly \& Associates.

\bibitem{c++}
{\sc Stroustrup, B.} 1986.
\newblock {\em The C++ Programming Language}.
\newblock Addison Wesley.

\bibitem{barret}
{\sc Wu, X., Bryant, B.R., Gray, J., Mernik, M.},
\newblock Component-based LR parsing.
\newblock {m Computer Languages, Systems \& Structures 36(1)}, 16-–33 (2010)

\bibitem{erik}
{\sc Schwerdfeger A, Van Wyk E.},
\newblock Verifiable Parse Table Composition for Deterministic Parsing
\newblock { Software Language Engineering, Lecture Notes in Computer Science  
Volume: 5969}    184-203  2010  

\bibitem{visser}
{\sc Bravenboer, M., Visser, E.},
\newblock Parse Table Composition.
\newblock { Software Language Engineering LNCS}, 74--94 (2009)

\bibitem{basten}
{\sc Basten, H.J.S., Vinju, J.J.},
\newblock Faster Ambiguity Detection by Grammar Filtering. 
\newblock Tenth Workshop on Language Descriptions, Tools, and Applications (LDTA 2010), Paphos, Cyprus, March 2010

%\bibitem{tatoo}
%{\sc Cervelle, J., Forax R., Roussel, G.}
%\newblock Tatoo: an innovative parser generator
%\newblock {\em Proceedings of the 4th international symposium on Principles and practice of programming in Java ACM}, 13-20 (2006)

\bibitem{c}
{\sc Ritchie, K.~.} 1988.
\newblock {\em \em The C Programming Language}.
\newblock Prentice Hall.

\bibitem{stroustrup}
{\sc Ellis, M.~A.} {\sc and} {\sc Stroustrup, B.} 1990.
\newblock {\em The Annotated C++ Reference Manual}.
\newblock Addison-Wesley.
            
%\bibitem{schrodingerstoken}
%{\sc Aycock, J.} {\sc and} {\sc Horspool, R.~N.} 2001.
%\newblock Schr\"{o}dinger's token.
%\newblock {\em Software, Practice \& Experience.\/}~{\em 31}, 803--814.

%0 Conference Paper
%1 1168057
%A Julien Cervelle
%A R\&\#233;mi Forax
%A Gilles Roussel 
%T Tatoo: an innovative parser generator
%B Proceedings of the 4th international symposium on Principles and practice of programming in Java
%@ 3-939352-05-5
%C Mannheim, Germany
%P 13-20
%D 2006
%R 10.1145/1168054.1168057
%I ACM 

\bibitem{aho2006}
{\sc Aho A.V., Lam, M., Sethi, R., Ullman, J.},
\newblock {\em Compilers: Principles, Techniques, and Tools (2nd Edition)}.
\newblock {Addison Wesley}, August 2006.

\bibitem{lgforte}
{\sc Rodr\'iguez-Le\'on, C. Garc\'ia, L.},
\newblock A Repository of LALR Conflictive Grammars 
\newblock {google-code: \tt \url{http://code.google.com/p/grammar-repository/}}, 2011.

%\bibitem{yagg}
%{\sc Coppit, D.} {\sc and} {\sc J., L.} 2005.
%\newblock yagg: an easy-to-use generator for structured test inputs.
%\newblock {\em 20th IEEE/ACM International Conference on Automated Software
%  Engineering November\/}, 356--359.

%\bibitem{yougen}
%{\sc Daniel~Malcolm, David Ly-Gagnon, P.~S.} {\sc and} {\sc Wang, H.-Y.} 2010.
%\newblock Grammar-based test generation with yougen.
%\newblock {\em Software, Practice \& Experience.\/}.
%\newblock [Online]. Available: {\tt
%  http://onlinelibrary.wiley.com/doi/10.1002/spe.1017/pdf}.

\bibitem{ford}
{\sc Ford, B.} 2002.
\newblock { Functional Pearl: Packrat Parsing: Simple, Powerful, Lazy,
  Linear Time.}
\newblock \\{\tt http://pdos.csail.mit.edu/~baford/packrat/icfp02/packrat-icfp02.pdf}.

%\bibitem{hudak}
%{\sc Hudak, P.} 1998.
%\newblock Modular domain specific languages and tools.
%\newblock In {\em ICSR '98: Proceedings of the 5th International Conference on
%  Software Reuse}. IEEE Computer Society, 134--142.

%\bibitem{koza}
%{\sc Koza, J.~R.} 1993.
%\newblock {\em Genetic programming: on the programming of computers by means of
%  natural selection}.
%\newblock MIT Press.

%\bibitem{maurer}
%{\sc Maurer, P.~M.} 1992.
%\newblock The design and implementation of a grammar-based data generator.
%\newblock {\em Software, Practice \& Experience.\/}~{\em 22}, 223--244.

%\bibitem{merrill}
%{\sc Merrill, G.~H.} 1993.
%\newblock Parsing non-LR( k ) grammars with yacc.
%\newblock {\em Software, Practice \& Experience.\/}~{\em 23}, 829--850.

%\bibitem{testlectrotest}
%{\sc Moertel, T.} 2004.
%\newblock {\em Test::LectroTest Manuals}.
%\newblock [Online]. Available: {\tt
%  http://search.cpan.org/dist/Test-LectroTest/}.

