One of the key contributions that made possible solving the PL-I
reserved-word versus identifier problem was 
the fact that the lexical analyzer can 
compute the \underline{exact} list of tokens 
expected by the syntax analyzer at any point of the scanning process.
This section describes the proposed  algorithm 
to compute this set. 
The algorithm uses simulation on the stack -
also known as symbolic interpretation - to achieve its goal.

The algorithm starts copying the parsing stack and callying the 
parser method $YYExpected$ with a reference to the parser stack:

%\IF {$i\geq maxval$} \STATE $i\gets 0$ \ELSE \IF {$i+k\leq maxval$} \STATE $i\gets i+k$ \ENDIF \ENDIF 
\begin{algorithmic}
\STATE    $YYExpected(parser\ stack)$
\end{algorithmic}

The pseudo-code for $YYExpected$ appears in Algorithm
\ref{algorithm:YYExpected}. In this and the next algorithms 
we will follow these conventions:
\begin{itemize}
\item $Q$ is the set of the states of the LR automaton
\item $\Sigma$ is the set of tokens or grammar terminals
\item $V$ is the set of variables or non-terminals
\item $P$ is the set of grammar productions
\item $ACTION$ denotes the LR action table: 
\[ ACTION : Q \times \Sigma \rightarrow Q \cup P \]
Where the kind of action is either a shift or a reduce:
\begin{eqnarray*}
ACTION(q, a) &=& shift\ q' \in Q\\
ACTION(q, a) &=& reduce\ by\ A \rightarrow \alpha \in P
\end{eqnarray*}

\item $GOTO$ is the LR automaton transition table restricted
to $V$:
\begin{eqnarray*}
GOTO : Q \times V &\rightarrow& Q\\
GOTO(q, A) &=& \ q' \in Q\\
\end{eqnarray*}
\end{itemize}

\begin{algorithm}[h] 
\caption{$YYExpected(\mathcal{S})$}
\label{algorithm:YYExpected}
\begin{algorithmic}%[1]
\REQUIRE {$Stack\ \mathcal{S}$}
\ENSURE The set of all expected tokens $\mathcal{E}$
\STATE $s = top(\mathcal{S})$
\STATE  $\mathcal{E} = \{ a \in \Sigma : \exists t \in Q\ such\ that\ ACTION(s, a) = shift\ t\ \}$
\STATE  $\mathcal{R} = \{  A \rightarrow \alpha \in P : ACTION(s, b) = reduce  A \rightarrow \alpha \}$
\IF {$\mathcal{R} \neq \emptyset$} 
  \STATE $\mathcal{E} = \mathcal{E} \cup YYSimStack(\mathcal{S}, \mathcal{R})$
\ENDIF
\RETURN $\mathcal{E}$
\end{algorithmic}
\end{algorithm}
Algorithm \ref{algorithm:YYExpected}  starts traversing the action table for the
state $s$ in the top of the stack. Tokens $a \in \Sigma$ 
for which there is a shift action 
are directly pushed in the set of expected 
tokens $\mathcal{E}$. 
The next step is to compute the set $\mathcal{R}$ of productions $A \rightarrow \alpha$  in
$P$ for which there
is a reduction. The remaining expected tokens
are computed calling $YYSimStack$, whose code appears in
Algorithm \ref{algorithm:YYSimStack}.


\begin{algorithm}[h]
\caption{$YYSimStack(\mathcal{S}, \mathcal{R})$}
\label{algorithm:YYSimStack}
\begin{algorithmic}[1]
\REQUIRE {$Stack\ \mathcal{S}, Productions\ \mathcal{R}$}
  \FORALL {$A \rightarrow \alpha \in \mathcal{R}$}
    \IF{$length(S) > length(\alpha)$}
      \STATE $Stack\ \mathfrak{S} = \mathcal{S}$
      \STATE \label{line:pop} $pop\ \mathfrak{S},\  length(\alpha)$
      \STATE $s = top(\mathfrak{S})$
      \STATE $n = GOTO(s, A)$
      \STATE $push\ \mathfrak{S}, n$
      \STATE \label{line:rec}$\mathcal{E} = \mathcal{E} \cup YYExpected(\mathfrak{S})$
    \ENDIF
 \ENDFOR
 \RETURN $\mathcal{E}$
\end{algorithmic}
\end{algorithm}

Algorithm \ref{algorithm:YYSimStack} 
consists of a simulation recursive process that 
branches at each available reduction.
The exploration is kept until no more 
productions can be applied for reduction.

The simulation mimics the LR parsing algorithm:
For each production $A \rightarrow \alpha \in \mathcal{R}$
the step of reducing by it is performed using a local stack 
$\mathfrak{S}$: As many states as the number of symbols in  the right hand side 
$\alpha$ are extracted from the stack (line \ref{line:pop}), leaving some state $s$ at the top.
The $GOTO$ table of the LR parser is then consulted to find which the 
next state $n$ is. The algorithm recursively calls back (line \ref{line:rec}) to 
$YYExpected$ %(Algorithm \ref{algorithm:YYExpected})
 to compute
the set of tokens expected in state $n$. The new tokens are added
to $\mathcal{E}$.
